A Hitchhiker's Guide to Reproducible Research" # replace with something serious

### 2021-05-04 manuscript.Rmd
add missing rocker citation close #24
master -> main closes #33

This tutorial is targeted at scholars with at least intermediate R knowledge and we assume that you have already installed R and RStudio (if not, check the `r link("https://github.com/aaronpeikert/repro-tutorial/blob/master/install.md", "installation guide")`). To create reproducible research projects in R, we have implemented our workflow in the  [`repro`-package](https://github.com/aaronpeikert/repro)[^repropackage]:

@@

Another advantage of Markdown is that you can easily version-control it with Git.
You may be familiar with Microsoft Words' "Track Changes"-Feature.
Version control is similar in its basic idea but more potent in its features.
Instead of tracking only one file, you can keep track of the whole project directory, and instead of losing all changes when you agree to them, the entire history is preserved transparently.
For example, this paper has been going through `r length(git2r::commits())` `r link("https://github.com/aaronpeikert/repro-tutorial/commits/master", "iterations")`.
To activate Git in a given project, you can call:
<!-- Hands-on: Introduce Git -->

@@

Note that this function is again composed of several functions, which you may inspect on `r link("https://github.com/aaronpeikert/repro-tutorial/blob/master/R/simulation.R", "GitHub")`.

@@

When you include code into an R Markdown, this code is run, and its results are included in the rendered document.
With R Markdown, you can write whole articles (`r link("https://github.com/aaronpeikert/repro-tutorial/blob/master/manuscript.Rmd", "like the one you are reading")`).
You could, in principle, even write large parts of the manuscript as part of your preregistration, i.e. the introduction, theoretical background, the method section and parts of the discussion.
If you have simulated data, you could even include fake results.
Such a dynamic document is arguably the most concrete preregistration possible.
One advantage is that writing the manuscript and writing the preregistration are no longer two distinct tasks.
Therefore a preregistration can be thought of as simply rearranging tasks that you have to do anyway.

@@

Of course, there a several ways to solve this problem.
One could build an R package to separate analysis and data or use code externalization in the RMarkdown to facilitating reuse or create a "master" script that combines all steps.
We found Make to be particularly useful for two reasons.
First, it is language agnostic, therefore fusing steps that are carried out in different programming languages.
Second, it does implement a simple yet powerful caching mechanism of results which is welcome in computing-intensive analysis often found in neuroimaging or machine learning.

@@

Modern data analysis is only feasible because we rely on much external software.
One downside is that a considerable number of external dependencies may hinder collaboration and long term reproducibility.
The critical problem is the fragility of a software setup.
Many components have to fit perfectly together, and each change, e.g. an update, risks that balance.
Updates are, of course, very much necessary.
Another problem is that the software environment is also variable and what works on one machine might not work on any other.
Docker circumvents such issues.
Instead of leaving it to the user which software should be installed how on their computer, Docker provides all software used in an analysis, down to the operating system.
Docker does this without interfering with the already installed software by using a virtual software environment that is independent of the host software environment.
Such a snapshot of the software stack is called "image".
Of course, packaging all needed software in such an image requires considerable amounts of storage.
However, two strategies help to keep the storage requirements manageable.
First, there is a community that maintains images for particular purposes.
For example, using R [rocker-citation].
Such images only contain what is necessary for a specific purpose.
Users can then install whatever they additionally need on top of these purpose-built images.
The image that was used for this article uses `r round(as.numeric(fs::file_size("reprotutorial.sif"))/1024^3, 2)`Gb of disc space.
It does include Ubuntu (a Linux operating system) R, RStudio, LaTeX as well as many R packages like the tidyverse (amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` packages.)
Second, instead of saving a binary image of the software, one can also save a textual description of the programmatic steps to install the software in a `Dockerfile`.
Such files are tiny (the `r link("https://github.com/aaronpeikert/repro-tutorial/blob/master/Dockerfile", "Dockerfile")` for this project is `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb big), but they rely on the assumption that all software repositories are available and contain the correct software version.
For archiving, you therefore have to store the image, as explained in [Archiving Docker].
The `repro`-package lets you specify required packages as part of the metadata of the RMarkdown:

@@

[^simhash]: https://github.com/aaronpeikert/repro-tutorial/blob/master/R/hash.R

>>>>

This tutorial is targeted at scholars with at least intermediate R knowledge and we assume that you have already installed R and RStudio (if not, check the `r link("https://github.com/aaronpeikert/repro-tutorial/blob/main/install.md", "installation guide")`). To create reproducible research projects in R, we have implemented our workflow in the  [`repro`-package](https://github.com/aaronpeikert/repro)[^repropackage]:

@@

Another advantage of Markdown is that you can easily version-control it with Git.
You may be familiar with Microsoft Words' "Track Changes"-Feature.
Version control is similar in its basic idea but more potent in its features.
Instead of tracking only one file, you can keep track of the whole project directory, and instead of losing all changes when you agree to them, the entire history is preserved transparently.
For example, this paper has been going through `r length(git2r::commits())` `r link("https://github.com/aaronpeikert/repro-tutorial/commits/main", "iterations")`.
To activate Git in a given project, you can call:
<!-- Hands-on: Introduce Git -->

@@

Note that this function is again composed of several functions, which you may inspect on `r link("https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation.R", "GitHub")`.

@@

When you include code into an R Markdown, this code is run, and its results are included in the rendered document.
With R Markdown, you can write whole articles (`r link("https://github.com/aaronpeikert/repro-tutorial/blob/main/manuscript.Rmd", "like the one you are reading")`).
You could, in principle, even write large parts of the manuscript as part of your preregistration, i.e. the introduction, theoretical background, the method section and parts of the discussion.
If you have simulated data, you could even include fake results.
Such a dynamic document is arguably the most concrete preregistration possible.
One advantage is that writing the manuscript and writing the preregistration are no longer two distinct tasks.
Therefore a preregistration can be thought of as simply rearranging tasks that you have to do anyway.

@@

Of course, there a several ways to solve this problem.
One could build an R package to separate analysis and data or use code externalization in the RMarkdown to facilitating reuse or create a "main" script that combines all steps.
We found Make to be particularly useful for two reasons.
First, it is language agnostic, therefore fusing steps that are carried out in different programming languages.
Second, it does implement a simple yet powerful caching mechanism of results which is welcome in computing-intensive analysis often found in neuroimaging or machine learning.

@@

Modern data analysis is only feasible because we rely on much external software.
One downside is that a considerable number of external dependencies may hinder collaboration and long term reproducibility.
The critical problem is the fragility of a software setup.
Many components have to fit perfectly together, and each change, e.g. an update, risks that balance.
Updates are, of course, very much necessary.
Another problem is that the software environment is also variable and what works on one machine might not work on any other.
Docker circumvents such issues.
Instead of leaving it to the user which software should be installed how on their computer, Docker provides all software used in an analysis, down to the operating system.
Docker does this without interfering with the already installed software by using a virtual software environment that is independent of the host software environment.
Such a snapshot of the software stack is called "image".
Of course, packaging all needed software in such an image requires considerable amounts of storage.
However, two strategies help to keep the storage requirements manageable.
First, there is a community that maintains images for particular purposes.
For example, using R [@boettigerIntroductionRockerDocker2017].
Such images only contain what is necessary for a specific purpose.
Users can then install whatever they additionally need on top of these purpose-built images.
The image that was used for this article uses `r round(as.numeric(fs::file_size("reprotutorial.sif"))/1024^3, 2)`Gb of disc space.
It does include Ubuntu (a Linux operating system) R, RStudio, LaTeX as well as many R packages like the tidyverse (amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` packages.)
Second, instead of saving a binary image of the software, one can also save a textual description of the programmatic steps to install the software in a `Dockerfile`.
Such files are tiny (the `r link("https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile", "Dockerfile")` for this project is `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb big), but they rely on the assumption that all software repositories are available and contain the correct software version.
For archiving, you therefore have to store the image, as explained in [Archiving Docker].
The `repro`-package lets you specify required packages as part of the metadata of the RMarkdown:

@@

[^simhash]: https://github.com/aaronpeikert/repro-tutorial/blob/main/R/hash.R
